---
title: "Clean rosters"
---

This notebook is to clean the main roster listed on the left part of [the PDF](data-raw/Club-Roster-Profiles-2024-01-01.pdf).

See [the home page](index.qmd) about using Pinpoint to get to this point.

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
```

## Import

This pulls in the first file that was exported from Pinpoint.

```{r}
roster_raw <- read_csv("data-raw/pinpoint/Roster.csv") |> clean_names()
```

## Roster cleaning

### Make short club name

I first need to build a club name of some type, which I'm doing from the file name. I'm basing the short code based on another dataset I have [rws-mls-salaries](https://github.com/utdata/rwd-mls-salaries).

I also move around the new column and remove the Pinpoint link variable no one can use anyway.

```{r}
add_short_name <- roster_raw |> 
  mutate(
    club_short = str_extract(file_name, "^\\w*") |> str_to_upper()
  ) |>
  relocate(club_short) |> 
  select(!c(file_name, validation_link))

add_short_name |> glimpse()
```

### Make a club name

```{r}
add_long_name <- add_short_name |> 
  mutate(
    club = case_match(
      club_short,
      "ATL" ~ "Atlanta United",
      "ATX" ~ "Austin FC",
      "CHI" ~ "Chicago Fire FC",
      "CIN" ~ "FC Cincinnati",
      "CLB" ~ "Columbus Crew",
      "CLT" ~ "Charlotte FC",
      "COL" ~ "Colorado Rapids",
      "DAL" ~ "FC Dallas",
      "DC" ~ "D.C. United",
      "HOU" ~ "Houston Dynamo FC",
      "LA" ~ "LA Galaxy",
      "LAFC" ~ "Los Angeles Football Club",
      "MIA" ~ "Inter Miami FC",
      "MIN" ~ "Minnesota United FC",
      "MTL" ~ "CF Montréal",
      "NE" ~ "New England Revolution",
      "NSH" ~ "Nashville FC",
      "NY" ~ "New York Red Bulls",
      "NYC" ~ "New York City FC",
      "ORL" ~ "Orlando City DC",
      "PHI" ~ "Philadelphia Union",
      "POR" ~ "Portland Timbers",
      "RSL" ~ "Real Salt Lake",
      "SEA" ~ "Seatle Sounders FC",
      "SJ" ~ "San Jose Earthquakes",
      "SKC" ~ "Sporting Kansas City",
      "STL" ~ "St. Louis City SC",
      "TOR" ~ "Toronto FC",
      "VAN" ~ "Vancouver Whitecaps FC",
      .default = NA
    )
  ) |> 
  relocate(club, .after = club_short)

add_long_name
```




### Make roster designation

The roster is really split into different categories with an overline header. That's bad data management so I create a `roster_type` variable to say what category they are in.

I have to do a lot of work to get the player categories right. You'd have to run the code line by line to really see it, but I annotated the lines below.

```{r}
roster_subtype <- add_long_name |> 
  mutate(
    roster_type = case_when( # <1>
      is.na(lag(club_short)) ~ "SENIOR ROSTER", # <2>
      lag(club_short) != club_short ~ "SENIOR ROSTER", #<3>
      str_detect(name, "[A-Z]{3}") ~ name, # <4>
      .default = NA
    )
  ) |>
  fill(roster_type) |> #<5>
  relocate(roster_type, .after = club) |> # <6>
  filter(!str_detect(name, "[A-Z]{3}")) # <7>

roster_subtype
```

1. Make a `roster_type` variable
1. I use `lag()` to basically find the first empty row which is the top row and set `roster_type` to "SENIOR ROSTER". This is just for the first row of the data. There are probably better ways to do this.
1. Here I find when a new row has a different team using `lag()`, and if so set that row to SENIOR ROSTER. This is a typical use of lag.
1. If the name variable has three capital letters in a row then it is one of the categories (always in all caps) so I use `name` value in `roster_type`. This is telling me when a new roster category is starting.
1. This `fill()` takes those roster categories and fills them down the blank rows until it hits another existing category. This gives every play their category.
1. I move the new column to closer to the front of the tibble.
1. I remove the category headers since we don't need those anymore.

### Supplemental spot name

Here we rename SUPPLEMENTAL SPOT 31 to just SUPPLEMENTAL SPOT. It was inconsistent and I feel like the "31" needs explanation anyway. (No, I dunno what it means yet, but it is the only number used.)

```{r}
roster_supp <- roster_subtype |> 
  mutate(
    roster_type = case_match(
      roster_type,
      "SUPPLEMENTAL SPOT 31" ~ "SUPPLEMENTAL SPOT",
      .default = roster_type
    )
  )

roster_supp |> count(roster_type)
```

### Stray pipe in value

[@philwest](https://x.com/philwest/status/1786748847055675643) noted there is a stray | character in one of the fields. I fix that here.

```{r}
roster_desig_fix <- roster_supp |> 
  mutate(
    roster_designation = str_remove(roster_designation, "\\|") |> str_trim()
  )
```



## Make a list of teams

When I later work on the "other" player designations, I need to see if I have all the teams represented in my files. I create a list of those teams here so I can use it later.

```{r}
clubs <- roster_desig_fix |> 
  distinct(club_short)

clubs
```

## Export the files

Saving our the csv file for posterity. The rds file is the same data but in R's native format. Also save out the clubs for later.

```{r}
roster_desig_fix |> write_csv("data-out/rosters.csv")
roster_desig_fix |> write_rds("data-processed/rosters.rds")
clubs |> write_rds("data-processed/clubs.rds")
```


