---
title: "Join into profiles"
---

This notebooks takes the two cleaned files and joins them together into a single file.

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
```

## Import

```{r}
rosters <- read_rds("data-processed/rosters.rds")
others <- read_rds("data-processed/others.rds")
```


## Joining the two lists

Now that the data has been cleaned properly, this isn't as huge a deal.

```{r}
roster_profiles_joined <- rosters |> 
  left_join(others, by = join_by(club_short, name))

roster_profiles_joined
```

And now to fill in missing false values. This is an interesting and complex maneuver.

Change all the variables where the datatype is logical, and use replace_na() to fill black valuse with FALSE.

```{r}
roster_profiles_clean <- roster_profiles_joined |> 
  mutate(
    across(where(is.logical), ~ replace_na(.x, FALSE) )
  )

roster_profiles_clean
```

## Export

```{r}
roster_profiles_clean |> write_csv("data-out/profiles.csv")
roster_profiles_clean |> write_rds("data-processed/profiles.rds")
```


