---
title: "Quality checks"
---

Because we are extracting this data from PDF's I used this file to troubleshoot and make sure all this is good.

## Fixed issues

- In the [Cleaning roster](01-cleaning-roster.qmd) notebook the `roster_type` == "SUPPLEMENTAL SPOT 31" was changed to just "SUPPLEMENTAL SPOT" for consistency. The roster has the "31" in all cases but seems odd to keep here. I can change later if needed.
- In `others` there were initially some missing players. I ended up piecing everything together in [Cleaning other](01-cleaning-other.qmd).
- Also in `others` I had to rework how players were awarded different types and notes because in the original list players can be listed more than once with different designations. I had to collapse all of that.

There aren't any known issues as of now, but I'm keeping this notebook around for now.

## Setup

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
library(janitor)
```


## Import the rosters

```{r}
rosters <- read_rds("data-processed/rosters.rds")

rosters |> glimpse()
```

Do I have all the teams?

```{r}
rosters |> 
  count(club_short)
```

There are [29 teams](https://en.wikipedia.org/wiki/Major_League_Soccer) in the MLS as of May 2, 2024.

Let's spot check some teams.

```{r}
rosters |> filter(club_short == "PHI")
```

## Others

We look at the others file here.

```{r}
others <- read_rds("data-processed/others.rds")

others |> glimpse()
```

### Missing teams

At one point we were missing teams. Here we make sure there are 29.

```{r}
others |> 
  count(club_short)
```

### Checking season-ending list

I noticed this because we were missing injured players. Here I check for them.

```{r}
others |> 
  filter(type_inj == TRUE)
```

## Profiles check

The last check of everything

```{r}
profiles <- read_rds("data-processed/profiles.rds")

profiles
```

Example for index page

```{r}
profiles |> filter(club_short == "ATX") |> 
 head(2) |> glimpse()
```

How many teams

```{r}
profiles |> 
  count(club_short)
```


